---
title: "Reducing Racial Disparities through Experience: Evidence from Medical Device Choice in Cardiac Catherization Labs"
subtitle: "Presenter: Chad Stecher, Arizona State University"
author: "Discussant: Ian McCarthy, Emory University and NBER"
date: "ASHEcon Annual Meeting, June 28, 2022"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{includepdf}
  - \usepackage{pgf}
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usepackage{pdftools}
  - \pgfplotsset{compat=1.15}
  - \usepackage{mathrsfs}
  - \usetikzlibrary{arrows}
  - \usetikzlibrary{patterns}  
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
knit: pagedown::chrome_print
---

<!-- Adjust some CSS code for font size, maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code, .remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(dplyr.summarise.inform = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD, ggrepel, xaringanExtra, webshot)
```


```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```


# Racial disparities in health care
- Lower life expectancy of more than 4 years for non-Hispanic Black Americans versus White Americans
- More than 6 years in 2020
- What can health care policy do to shrink this gap?
  - Outright discrimination
  - *Implicit bias* (where this paper fits?)
  
---
# This paper
*Main Question*: Does experience with Black patients reduce the racial gap in DES use?

<br>
*Approach*: LPM regression of DES use (separately by race) against different measures of experience

<br>
Unique data on team construction!


---
class: inverse, center, middle
name: questions

# Some high-level thoughts

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Learning differentially by race?
- Other evidence that physicians learn differentially by the race of the patient?
- Consistent with implicit bias, but haven't seen in a learning context
  - Maybe related to Sarsons, **Interpreting Signals in the Labor Market: Evidence from Medical Referrals**
- Role of implicit bias could work in different directions, depending on strength of priors and early experience

---
# Who is learning?
- Hard to think about team learning in this context
- Specialist seems to make the decision of whether to use DES