---
title: "Peer Participation in the Medicare Shared Savings Program"
subtitle: "Presenter: Youcheng Hou"
author: "Discussant: Ian McCarthy"
date: "ASHEcon Annual Meeting, June 28, 2022"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{includepdf}
  - \usepackage{pgf}
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usepackage{pdftools}
  - \pgfplotsset{compat=1.15}
  - \usepackage{mathrsfs}
  - \usetikzlibrary{arrows}
  - \usetikzlibrary{patterns}  
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
#knit: pagedown::chrome_print
---

<!-- Adjust some CSS code for font size, maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code, .remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(dplyr.summarise.inform = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD, ggrepel, xaringanExtra, webshot)
```


```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```


# Physician networks and voluntary participation
- Lots of work measuring static physician networks
- Very little work studying the effects of such networks
  - Even less work on the formation and evolution of those networks (more on this later)
- Similarly very little work on voluntary program participation in health care
  - Look into "Selection and Causal Effects under Voluntary Programs" by Gupta et al.

---
# This paper
*Main Question*: Does my peers' participation in a shared savings program affect my own participation?

<br>
*Approach*: LPM regression of my participation against a weighted average of my lagged peers' participation, instrumented using lagged peers-of-peers

<br>
A ton of impressive data work!

---
# Some descriptive work
- Would love to see some graphs describing the networks, especially the instrument (peer-of-peers)


---
class: inverse, center, middle
name: questions

# Some thoughts and questions

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Reflection
- Are lagged values necessary?
- Already using leave-one-out peer average
- Ahn and Trogdon (2017) use contemporaneous (leave-one-out) peer average with lagged peer-of-peer average as instrument

---
# Weighting
- Results much larger when not weighting peer averages by patient counts
- Could instead be a reflection of measurement error in the peer group (Angrist, 2014)

---
# Mechanisms
- HHI highly correlated with peer group structure
- Is recruitment effect identified separately from "social learning"?
- Recruitment is a two-way street...more likely to recruit if program is running successfully
- Suggests that, as I learn about my peers' performance, recruitment may be harder or easier
  - Is there some way to exploit "successfull" versus "unsuccessful" ACOs here?

---
# Network formation
- Is my peer-of-peer exogenous?
  - If the existence of an edge is endogenous, then the absence of an edge should be endogenous too
- Analysis allows for time varying peer groups, but what assumptions are needed on the evolution of the network
- Given the need for sufficiently large ACOs, practices must be forward-thinking
  - Could you exploit system-level information here? Plausibly exogenous shock to ACO participation

