---
title: "Healthy Reviews! Online Physician Ratings Reduce Healthcare Interruptions"
subtitle: "Presenter: Danny R. Hughes, Georgia Tech"
author: "Discussant: Ian McCarthy, Emory University and NBER"
date: "ASHEcon Annual Meeting, June 29, 2022"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{includepdf}
  - \usepackage{pgf}
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usepackage{pdftools}
  - \pgfplotsset{compat=1.15}
  - \usepackage{mathrsfs}
  - \usetikzlibrary{arrows}
  - \usetikzlibrary{patterns}  
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
knit: pagedown::chrome_print
---

<!-- Adjust some CSS code for font size, maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code, .remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(dplyr.summarise.inform = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD, ggrepel, xaringanExtra, webshot)
```


```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```

```{r include=FALSE}
finding.phy.mcare <- tribble(
  ~"Difficulty", ~"Yr2008", ~"Yr2009", ~"Yr2016", ~"Yr2017",
  "No Problem", 71, 78, 64, 69,
  "Small Problem", 10, 10, 15, 13,
  "Big Problem", 18, 12, 20, 14
)

finding.phy.comm <- tribble(
  ~"Difficulty", ~"Yr2008", ~"Yr2009", ~"Yr2016", ~"Yr2017",
  "No Problem", 72, 71, 63, 59, 
  "Small Problem", 13, 8, 16, 18, 
  "Big Problem", 13, 21, 20, 22
)

looking.pcp <- tribble(
  ~"Insurance", ~"Yr2008", ~"Yr2009", ~"Yr2016", ~"Yr2017",
  "Medicare and MA", 6, 6, 8, 9, 
  "Commercial", 7, 8, 10, 11
)

```

# Consumer search
- Highly relevant and underappreciated structure for studying health care decisions
- Basic idea:
  - Patients incur a cost to accessing/assimilating information about physicians
  - Patients have some beliefs about their utilities over all possible options
  - Patients stop searching when costs meet or exceed expected benefits

---
# Context
- Physicians retire and patients must find a new physician
- Online reviews lower the cost of this search and increase probability of a more recent visit to a new physician (not sure that the first implies the second, but more on this later)

---
# Looking for New PCP

.center[
```{r echo=FALSE}
pcp <- pivot_longer(looking.pcp,
                           cols = starts_with("Yr"),
                           names_to = "Year",
                           names_prefix = "Yr",
                           values_to="Percent")
ggplot(pcp, aes(x=Year, y=Percent, fill=Insurance, label=paste0(sprintf("%1.1f", Percent),"%"))) +
  geom_bar(stat="identity") +
  geom_text(position=position_stack(vjust=0.5), color="white", size=3) +
  scale_fill_grey() +
  theme_bw()
```
]


MedPAC *Report to Congress*, March 2012 and March 2020


---
# Finding a PCP among Medicare and MA

.center[
```{r echo=FALSE}
mcare.pcp <- pivot_longer(finding.phy.mcare,
                           cols = starts_with("Yr"),
                           names_to = "Year",
                           names_prefix = "Yr",
                           values_to="Percent") %>%
  mutate(order=case_when(
    Difficulty=='Big Problem' ~ 2,
    Difficulty=='Small Problem' ~ 1,
    Difficulty=='No Problem' ~ 0
  ))
ggplot(mcare.pcp, aes(x=Year, y=Percent, fill=forcats::fct_reorder(Difficulty, order), label=paste0(sprintf("%1.1f", Percent),"%"))) +
  geom_bar(stat="identity") +
  geom_text(position=position_stack(vjust=0.5), color="white", size=3) +
  scale_fill_grey() +
  labs(fill="Difficulty") +
  theme_bw()
```
]

MedPAC *Report to Congress*, March 2012 and March 2020


---
# Finding a PCP among Commercial Insurance

.center[
```{r echo=FALSE}
comm.pcp <- pivot_longer(finding.phy.comm,
                           cols = starts_with("Yr"),
                           names_to = "Year",
                           names_prefix = "Yr",
                           values_to="Percent") %>%
  mutate(order=case_when(
    Difficulty=='Big Problem' ~ 2,
    Difficulty=='Small Problem' ~ 1,
    Difficulty=='No Problem' ~ 0
  ))
ggplot(comm.pcp, aes(x=Year, y=Percent, fill=forcats::fct_reorder(Difficulty, order), label=paste0(sprintf("%1.1f", Percent),"%"))) +
  geom_bar(stat="identity") +
  geom_text(position=position_stack(vjust=0.5), color="white", size=3) +
  scale_fill_grey() +
  labs(fill="Difficulty") +
  theme_bw()
```
]

MedPAC *Report to Congress*, March 2012 and March 2020


---
# Approach
- Research design: DD
- Sample: Patients of retired physicians in two periods
- Pre period: 2007-2010
- Post period: 2015-2018
- Control (Treatment): 18 (16) top 100 population cities, based on cities with fewest (most) Yelp reviews per capita and nonmissing reviews in pre and post period

---
class: inverse, center, middle
name: questions

# Some thoughts and questions

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Practical motivation
- Do people use Yelp to find new PCPs?
- You write, "...a relatively small proportion of patients rely on these platforms for choosing their physicians."
- But some work **does** suggest that patients value online reviews in health care


---
# Theoretical motivation
- Really like the idea, but...
- Paper considers search as decision to find a physician today or not
- Are patients "searching" over physicians, or are they just deciding whether to take an action on a given day?
- Is $v$ independent of $E[Days]$? How?
- Searching for a physician on a given day is not the same as visiting the physician that day
  - Search could *increase* time to visit because patients are drawn to the most capacity-constrained physicians
  - How should I think of waiting time here?
  
  
---
# General concern
- Treated cities very different than control cities (e.g., Laredo vs Anaheim), probably not just in levels
- Could relate to practice sizes and consolidation, thereby affecting within-practice referrals
- Null results (using days) when including retirement year/quarter FEs

---
# Identification and estimation
- Is 2x2 DD appropriate here? 
  - Treatment/control is continuous (availability of online reviews)
- IV using other Yelp reviews
  - Does this satisfy exclusion restriction? Only conditional on overall use of Yelp
  - Not if early adoption of Yelp (in general) is correlated with unobservables or if Yelp causes additional retirements
- Are cities the right level of treatment/control?
  - Laredo is isolated, Anaheim is right next to LA
  
---
# Some potential considerations...
1. More (or fewer) details on model
2. Treatment at county or HRR-level instead of city/zip
3. Measure visits at practice-level instead of physician-level
4. Visualizations of physician markets over time between treatment/control (total retirements, number of physicians, number of practices, size of practices, etc.)

Bonus question: Why do physician's retire and where do they retire most?


