---
title: "The Effect of Physician Group Mergers on the Health Outcomes of Medicare Beneficiaries"
subtitle: "Presenter: Brett Wendling"
author: "Discussant: Ian McCarthy"
date: "SHESG Annual Meeting, October 8, 2021"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{includepdf}
  - \usepackage{pgf}
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usepackage{pdftools}
  - \pgfplotsset{compat=1.15}
  - \usepackage{mathrsfs}
  - \usetikzlibrary{arrows}
  - \usetikzlibrary{patterns}  
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
#knit: pagedown::chrome_print
---

<!-- Adjust some CSS code for font size, maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code, .remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(dplyr.summarise.inform = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD, ggrepel, xaringanExtra, webshot)
```


```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```


```{r include=FALSE}
phy.status <- tribble(
  ~"Status", ~"Yr2012", ~"Yr2014", ~"Yr2016", ~"Yr2018",
  "Owner", 53.2, 50.8, 47.1, 45.9,
  "Employee", 41.8, 43.0, 47.1, 47.4,
  "Independent", 5.0, 6.2, 5.9, 6.7
)

phy.practice <- tribble(
  ~"Type", ~"Yr2012", ~"Yr2014", ~"Yr2016", ~"Yr2018",
  "Solo", 18.4, 17.1, 16.5, 14.8,
  "Single Specialty", 45.5, 42.2, 42.8, 42.8,
  "Multi Specialty", 22.1, 24.7, 24.6, 25.2,
  "Hospital", 5.6, 7.2, 7.4, 8.0,
  "Faculty", 2.7, 2.8, 3.1, 3.0,
  "Other", 5.7, 5.9, 5.7, 6.2
)

phy.size <- tribble(
  ~"Size", ~"Yr2012", ~"Yr2014", ~"Yr2016", ~"Yr2018",
  "Less than 5", 40.0, 40.9, 37.9, 35.7,
  "5 to 10", 21.4, 19.8, 19.9, 20.8,
  "11 to 24", 13.4, 12.1, 13.3, 12.7,
  "25 to 49", 7.1, 6.3, 7.4, 7.6,
  "50 or more", 12.2, 13.5, 13.8, 14.7
)
```

---
class: inverse, center, middle
name: background

# Important question!

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Changing physician landscape
.center[
```{r echo=FALSE}
phy.status <- pivot_longer(phy.status,
                           cols = starts_with("Yr"),
                           names_to = "Year",
                           names_prefix = "Yr",
                           values_to="Percent")
ggplot(phy.status, aes(x=Year, y=Percent, fill=Status, label=paste0(sprintf("%1.1f", Percent),"%"))) +
  geom_bar(stat="identity") +
  geom_text(position=position_stack(vjust=0.5), color="white", size=3) +
  scale_fill_grey() +
  theme_bw()
```
]

Data from AMA Physician Practice Benchmark Survey, 2012-2018

---
count: false
# Changing physician landscape

.center[
```{r echo=FALSE}
phy.practice <- pivot_longer(phy.practice,
                           cols = starts_with("Yr"),
                           names_to = "Year",
                           names_prefix = "Yr",
                           values_to="Percent") %>%
  arrange(Percent)
ggplot(phy.practice, aes(x=Year, y=Percent, fill=factor(Type, levels=c("Other","Faculty","Hospital","Multi Specialty", "Single Specialty", "Solo")), label=paste0(sprintf("%1.1f", Percent),"%"))) +
  geom_bar(stat="identity") +
  labs(fill="Practice Type") +
  geom_text(position=position_stack(vjust=0.5), color="white", size=3) +
  scale_fill_grey() +
  theme_bw()
```
]

Data from AMA Physician Practice Benchmark Survey, 2012-2018


---
count: false
# Changing physician landscape

.center[
```{r echo=FALSE}
phy.size <- pivot_longer(phy.size,
                           cols = starts_with("Yr"),
                           names_to = "Year",
                           names_prefix = "Yr",
                           values_to="Percent") %>%
  arrange(Percent)
ggplot(phy.size, aes(x=Year, y=Percent, fill=factor(Size, levels=c("50 or more","25 to 49","11 to 24","5 to 10","Less than 5")), label=paste0(sprintf("%1.1f", Percent),"%"))) +
  geom_bar(stat="identity") +
  labs(fill="Practice Size") +
  geom_text(position=position_stack(vjust=0.5), color="white", size=3) +
  scale_fill_grey() +
  theme_bw()
```
]

Data from AMA Physician Practice Benchmark Survey, 2012-2018

---
count: false

# Changing Physician Landscape

Physicians are now:
- less likely to own their own practice
- more likely to be employed by a hospital or part of practice owned by hospital
- more likely to be part of a very large practice

--

**Relatively little research on practice structure**

---
class: inverse, center, middle
name: review

# What does this paper do?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# In pictures

.pull-left[
**Practice at time** $t$
<br>
<br>
<br>
<br>
<br>
<br>
<br>
]

.pull-right[
**Practice at time** $t+1$
<br>
<br>
<br>
<br>
<br>
<br>
<br>
]

--

.pull-left[
at least half of physicians leave practice ("merger-to")
]

.pull-right[
at least 10% new physicians ("merger-from")
]

---
# Estimation and Identification

- Outcomes: Mortality and progression of disease
- Identification: Conditional independence assumption
- Estimation: OLS with indicators for different merger types and controls for patient characteristics


---
class: inverse, center, middle
name: questions

# Some thoughts and questions

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Capturing mergers in the data

- What happens if 50% of physicians leave a practice, and the practice subsequently (or concurrently) gains more than 10% of new physicians?
-

---
# Are these mergers?

.pull-left[
**Sure**
- Bulk of a practice dissolves (50%)
- Existing practice expands by some meaningful amount (10%)
]

.pull-right[
**Not really**
- "merger-to" practices still exist
- very hard to distinguish horizontal vs vertical
]

---
# If not mergers, what?

- Disruption in care
- Information transmission within practices
- Variations in practice styles due to peers


---
# Identification

- Physicians may change practices due to their current patient profile or restrictions on how they can practice
- Sample selection? How often does a person leave their physician after the physician changes practices?
- Block estimation predicts treatment using patient characteristics, but nothing about physician practices. What source of endogeneity is this intended to adjust for?
- In general, no physician characteristics used in the analysis?

---
# Estimation

- What is the estimand of interest?
- Not an ATT without strong homogeneity assumptions (across people and over time)
- Staggered treatment, etc.

---
# GIF expressing remorse for bringing this up

---
# Final ideas

- Thinking about movement across practices...
- Could exploit variation in non-compete agreement enforcement (Hausman and Lavetti, 2019)
